Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-11-18 Latex_files/Thesis/chapters/chapter 4/main.tex
updates to chapter 4, among others

After computing its $T$-matrix via \textit{mstm} and supplying that 
to \textit{ott} we compute the optical force exerted by a $50\ mW$ laser
via \eqref{eq:optical_force} in the axial direction while the dimer is 
in its 'standard' orientation. As expected we see a single point where 
the dimer will be in equilibrium, the linear fit in fig.~\ref{fig:paradigmatic}
(a) indicates a that the force can be modelled as a harmonic potential close 
to the equilibrium position ($F_z\approx-\kappa_z z$). The second point 
where the axial force goes to 0 cannot be considered as equilibrium position 
as the positive gradient indicates that the trap is unstable unless Brownian 
motion is ignored.

@@

We can see that both equilibrium positions have comparable axial trap 
stiffness ($\kappa_z$), however the difference in the transverse trap 
stiffness ($\kappa_x$) is far more noticeable. The same dimer was trapped
at each of the axial equilibrium positions and the transverse force was 
evaluated. While in all three cases the dimer can be trapped the linear
range where that would typically associated with a stable trap is far 
narrower in the 'standard' orientation compared to the 'inverted' cases. 
This highlights one of the challenges involved with studying 
asymmetric particles, even though its a simple enough process to trap them 
they maybe characterised very differently depending on their relative position 
and orientation towards the focus. This can have a significant impact on 
rheological studies - or attempting to probe any local property - as the 
variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's centre of mass spheres, 
		where a positive force indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation parameters are used here as in fig~\ref{fig:paradigmatic}(a) and (c). The blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

@@

\newpage
\subsection{Non-trivial equilibrium configurations}
\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned with the electric 
field is relatively simple as the orientational torque is minimised (see
Eq.\ref{eq:opt_torque}). Meaning once trapped the dimer is unlikely to change
orientation enough to escape the trap. Regardless, that does not rule out the 
possibility that there is a stable configuration where the orientation not 
strictly vertical, in fact most experimental work with symmetric nano-dimers 
will trap them lying perpendicular to the beam direction \cite{Ahn2018, 
Reimann2018}. Unlike in Sec.~\ref{sec:eq_positions} we cannot simply measure 
the optical force and torque as the parameter space is too large and determining 
if a particular position and orientation is stable is not clear based solely 
on force and torque measurements \cite{Bui2017}. Using the same simulation 
parameters as before we ran a number short simulations (total simulation time 
was $0.005$ s) with the laser power increased to $500\ mW$. Each simulation 
started with the dimer in a different starting position and orientation, due 
to the high laser power the dimers either escaped the trap or were stably 
trapped. The $z-\theta$ phase space - where $\theta$ is the angle between the 
direction of beam propagation and the dimer's orientation vector ($\theta=0^
\circ$ is the 'standard' orientation) - can be divided into different regions 
depending on which equilibrium configuration is reached.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
		\caption{}
	\end{subfigure}
	%\captionsetup{hangindent=2.35cm}
	\caption{Map of $z-\theta$ phase space using a dimer of size ratio 2 with 
		a laser power of 500 mW ($\theta=0\circ$ is the 'standard' orientation 
		and $\theta=180^\circ$ is the 'inverted' orientation). The stable 
		configurations are indicated by the larger circles and the starting 
		conditions are colour coded to match the stable point they end up in. 
		Right hand render shows a dimer in its off-axis configuration.}
	\label{fig:off_axis}
\end{figure}

Interestingly while the trap strength of these off-axis traps are similar 
in magnitude to the vertically aligned dimers, but when the laser power 
is lowered (around $5\ mW$) the traps become metastable; after reaching its'
equilibrium configuration the particle behaves similarly to a typically
trapped dimer but the trapping potential is small enough that the dimer
escapes in as little time as less than a second or after nearly a full 
3 seconds. Running similar simulations but for vertical configurations 
sees the dimer remaining trapped, even after 30 seconds of run time, 
indicating that the trapping potential is far greater than the thermal 
energy. This suggests that the reason this off-axis configuration is 
metastable is due to the increased rotational freedom. Similar 
configurations have been explored with ellipsoids; Zhu \textit{et al.} 
looked at the dynamics of various elliptical particles and found that 
regardless of shape or initial orientation the particle would tend towards
either a purely vertical or purely horizontal orientation \cite{Zhu2021}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Continuous rotational motion due to second-order scattering}
One aspect that has yet to be covered in depth with regards to spherical 
aggregates of any construction is their interaction with circularly 
polarised light. For homogenous spheres the optical torque is regarded as
being negligible as the spin density cannot impart angular momentum while 
propagating in a homogenous medium. Dimers however, have been shown to
experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
while trapped in a circular polarised beam. In our simulations we found 
that dimers would rotate about their long axis when trapped in circularly
polarised light. In this section we want to discuss how this behaviour 
is influenced by size, position, and orientation; and furthermore, we wish
to address possible explanations for this behaviour, as none of the current
theories into optically induced rotation seem plausible.

\subsection{Polarisation Dependency on Dimer trajectory}
First we wanted to confirm that the rotation seen in our simulations is
based on the polarisation of the trapping beam. To that end, we simulated 
the motion of an optically trapped dimer in beams of varying polarisation 
($NA=1.2$, $P=100\ mW$), the dimer is composed of polystyrene ($n_p=1.59$, 
$n_m=1.33$). Each simulation was run for $1$ second ($\Delta t =10^-5$) 
and at the end we looked at the orientational time series; the dimer's 
orientation is recorded as a quaternion which can be easily converted to 
a 3-dimensional rotation matrix. By considering only the transverse 
components ($U_{1,x}$, $U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation 
matrix and taking the Fourier transformation of their time series reveals 
the rotational frequency. The laser power is set to $100\ mW$ to avoid 
large thermal fluctuations and so that the Fourier series of the transverse 
components approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function 
centred at the rotational frequency $\omega_{rot}$.

@@

This shows us that these optical rotations are polarisation dependent 
and not merely an example of the dimer scattering light asymmetrically. 
The question then becomes, by what mechanism is the angular momentum of
the trapping beam being transferred to the dimer.

The rotation was first noted by Vigilante and co-workers who only 
considered this behaviour for a symmetric dimer \cite{Vigilante2020}.
In their work they attribute this to spin-curl effects, in which the 
curl of the spin density leads to a second order optical force that 
orbits around the beams central axis \cite{Yevick2017}. While several 
papers have demonstrated this phenomena \cite{Zhao2007,Zhao2009, Wang2010} 
it was only properly formalised by by \cite{Ruffner2012}. In which they 
showed that the seemingly random trajectory of a trapped sphere was biased
by the polarisation state of the trapping beam. While not immediately 
evident from the trajectory the helicity of the trapping beam was revealed 
by computing the particle's probability flux using.
\begin{align}
	j(r) = \frac{1}{N-1} \sum_{j=1}^{N-1}
	\frac{r_{j+1}-r_j}{\tau}\delta_{sigma_j}\left(r-\frac{r_{j+1}+r_j}{2}\right)
	\label{eq:prob_flux}
\end{align}

@@

While the results from \cite{Ruffner2012} suggest that the 
optical rotation seen in asymmetric dimers is attributed to 
the same spin-curl forces there are several discrepancies that 
cannot be explained purely by the spin-curl force. Firstly, 
there is the question of how the spin-curl force results in an 
optical torque: Comparing figures~\ref{fig:Ruffner_Grier}(c) 
and (e) we can see that the behaviour of the 7 sphere cluster 
rotates about the central axis whereas the single sphere has 
an only slight orbital bias in its motion. The authors attributed 
this to the fact that the cluster is scattering light where the 
spin-curl force is more substantial due to it having a wider 
profile. This would suggest that particles' whose longer axis 
lies in the plane perpendicular to the direction of propagation 
should experience any notable degree of torque. And indeed this 
appears to be the case for a number of experiments involving 
nano-dimers and ellipsoidal particles \cite{Ahn2018, Reimann2018}.

\subsection{Comparison of optical torques for Sphere and Dimers}
It does raise the question of how come our simulations show that 
despite the dimer being orientated so their long axis' align with 
the parallel to the direction of propagation they still readily 
rotate. You would not expect that the addition of a single additional 
sphere should drastically adjust the torque especially if said sphere 
is relatively small. However when we measured the optical torque 
of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we found the 
exact opposite. In both cases we used the same trapping beam as used 
for fig.~\ref{fig:rotation_vs_pol} but with a circularly polarised 
beam. Both the sphere and dimer were rotated in the $x-z$ plane and 
the all three components of the optical torque were recorded.
\begin{figure}[h!]
	\centering	
	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
	\caption{Optical toque experienced by a dimer ($a_{I}/a_{II}=10$) and a 
	single isotropic sphere. Both were rotated in the $x-z$ plane and the angle 
	between $U_z$ and the beam axis gives the polar angle.The solid lines denote 
	the torque experienced by the dimer whereas the dashed lines represent the 
	torque experienced by the sphere.}
	\label{fig:sphere_dimer_torque}
\end{figure}

The torques about the $x$ and $y$ axis can be somewhat understood 
as the second sphere is being drawn back towards the centre of the 
trap by the gradient forces. The same cannot be said for the $z$-
component of the optical torque; while this effect could still be 
attributed to the spin-curl force, but it is clear that the internal 
scattering between the two spheres has some unintended effects.

\subsection{Rotational frequency as a function of size ratio and orientation}
Next we want to see how the rotational speed is influenced by 
different sized dimers. Intuitively, you would expect that a 
larger would experience a greater torque and therefore rotate 
faster. By repeating the same kinds of simulation as used in 
\ref{fig:rotation_vs_pol} but for a circularly polarised beam 
$\phi=90\circ$ it was found that not only is the rotation rate 
dependent on the size of the dimer, but also on its orientation 
and therefore their axial position.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
  \caption{Rotation rate plotted against dimer size ratio while trapped in a
  	circularly polarised beam; a positive rotation rate indicates clockwise 
  	rotation, whereas a negative rotation rate indicates counter-clockwise 
  	rotation. The red line is for the case of a dimer in its 'standard' 
  	orientation. The blue line is for the case when the dimer is in its 
  	'inverted' orientation while trapped above the focus of the beam. And 
  	lastly the green line is for the case when the dimer is in its 
  	'inverted' orientation, but when it is trapped below the focus of the 
  	beam.}
\end{figure}

It is difficult to see from the graph, but the rotation rate never
truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which would
imply that a second sphere of radius $200\,{\rm nm}$ is enough to
induce rotational motion. This brings into question what mechanism is
generating the optical torque. We used \textit{mstm} to look at the stokes
parameters from the scattered field from a simple plane wave incident
on a symmetric dimer, the proportion of circularly polarised light is 
minimal compared to the proportion of plane polarised light, which 
indicates that this rotational motion is not due to any inhomogeneity 
in the dimer that might impart angular momentum to the scattered beam - 
as compared to a anisotropic scatterer like vaterite.

These results are somewhat contradictory to other work with silica dimers
\cite{Ahn2018, Debuysschere2023,Reimann2018}; previous experiments
have trapped the dimer in an orientation perpendicular to the beam
propagation direction. The rotational motion is attributed to the
dimers asymmetric geometry creating an unbalanced polarisation susceptibility
along its long axis as compared to its short axis; therefore its long
axis is aligned with the polarisation vector and can rotate
freely \cite{Ahn2018}. This however cannot be the case with our
simulations as the dimer rotates about its long axis, meaning there
cannot be an asymmetric axis to align with the beam's polarisation
vector. Furthermore, we see a non-linear increase in the rotational
speed of our dimers with size, the drag torque from the surrounding
fluid is $\propto r^3$ so the expectation is that the rotation
frequency should fall off with increasing size. This indicates that
the rotational motion is due to the shape asymmetry of the dimer and
not solely due to the beam's angular momentum. Measurement of this
photo-kinetic force is difficult to achieve due to the fact that
previous analysis was conducted in the Rayleigh regime, where the
polarizability of our dimer can be approximated as:
\begin{align}
  \bold{p}({\bf r},t)
  =
  \alpha_x E_x({\bf r},t)\hat{\bf e}_x
  + \alpha_yE_y({\bf r},t)\hat{\bf e}_y
  + \alpha_zE_z({\bf r},t)\hat{\bf e}_z
\end{align}

where the polarizability is given as a 3D vector for the three
principle Cartesian directions. In order to measure the magnitude of
second order contributions we would need to construct a dipole array
that fully captures the scattering of a dimer. Measuring the optical 
torque makes it clear that the polarizability is a contributing factor 
to this optical rotation phenomena.

\subsection{Gyroscopic Precession using asymmetric dimers}
As mentioned in section~\ref{sec:off-axis} for specificity sized dimers
there is the potential for non-vertical trapping orientations in which 
the dimer is still stably trapped. When a circularly polarised beam is 
used the dimer exhibits gyroscopic precession. As shown in fig.
~\ref{fig:gyro} the dimer's trajectory exhibits periodic rotation about 
its long axis $U_z$, precession motion as the dimer spins around the 
central beam axis, and nutation motion as the dimer's orientation rocks 
backwards and forwards. 
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
		off axis orientation with a circularly polarised beam ($P= 100\ mW$). 
		The far left column depicts the dimer's centre of mass position with 
		time; middle two columns are the $x$, $y$, and $z$ components of the 
		vectors $u_x$ and $u_y$; last column depicts the components of the 
		vector $u_z$ which defines the dimer's orientation.}
	\label{fig:gyro}
\end{figure}

Applying a Fourier analysis to the above trajectory reveals the 3 fundamental
frequencies typically associated with precession; the $u_{z,1}$ and $x(t)$ 
series show a precession frequency of $~1.33\ Hz$ whereas the series $u_{x,1}$ 
and $u_{y,1}$ show a combined periodic signal - a rotational frequency of 
$~23\ Hz$ and a nutation frequency of $~20\ Hz$. Previous studies into 
amorphous silica nanoparticles found a linear relationship with the 
rotational frequency and the laser power, but no such relationship existed 
with the precession frequency.

This gyroscopic motion has been demonstrated previously in nanoparticles 
\cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} but has not been observed 
for micron scale aggregates. Since the torque applied to the dimer is 
computed by evaluating the beam coefficients of the scattered field it is 
difficult to apply this result to micro-rheology experiments as one would 
need to know the exact magnitude of the optical torque ahead of time in 
order to make estimations about the local fluid viscosity. This is trivial 
for a birefringent spherical particle, less so for spherical aggregates 
whose equilibrium position and orientation are unknown. However, further 
analysis of the mechanism behind the precessive motion of off-axis dimers
may provide insights into controlling Brownian motion. An experimental work
trying to 'cool' nano-dimers by controlling the motion in all 6 degrees of 
freedom found that even while the rotation about the short axis' could be 
controlled the free rotation around the dimers' long axis resulted in an 
unpredictable torsional vibration \cite{Bang2020}. Understanding how 
rotational motion arises in the Mie-Regieme could allow researchers to 
build a robust theoretical framework to construct beam structures that 
eliminate any unwanted rotational motion from a target particle. Conversely, 
the same framework could allow for precise measurements of the optical 
torque applied to a target particle, allowing for characterisation of 
complex shaped particles' interactions with an optical trap.

>>>>

After computing its $T$-matrix via \textit{mstm} and 
supplying that to \textit{ott} we compute the optical 
force exerted by a $50\ mW$ laser via \eqref{eq:optical_force} 
in the axial direction while the dimer is in its 'standard' 
orientation. As expected we see a single point where 
the dimer will be in equilibrium, the linear fit in 
fig.~\ref{fig:paradigmatic}(a) indicates a that the 
force can be modelled as a harmonic potential close to 
the equilibrium position ($F_z\approx-\kappa_z z$). 
The second point where the axial force goes to 0 cannot 
be considered as equilibrium position as the positive 
gradient indicates that the trap is unstable unless 
Brownian motion is ignored.

@@

We can see that both equilibrium positions have comparable 
axial trap stiffness ($\kappa_z$), however the difference 
in the transverse trap stiffness ($\kappa_x$) is far more 
noticeable. The same dimer was trapped at each of the axial 
equilibrium positions and the transverse force was evaluated. 
While in all three cases the dimer can be trapped the linear
range where that would typically associated with a stable 
trap is far narrower in the 'standard' orientation compared 
to the 'inverted' cases. This highlights one of the challenges 
involved with studying asymmetric particles, even though its 
a simple enough process to trap them they maybe characterised 
very differently depending on their relative position and 
orientation towards the focus. This can have a significant 
impact on rheological studies - or attempting to probe any 
local property - as the variance in trap strength can result 
in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's 
		centre of mass spheres, where a positive force 
		indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation 
		parameters are used here as in fig~\ref{fig:paradigmatic}
		(a) and (c). The blue curve representing the force 
		response for a dimer in its standard orientation, orange 
		being the inverted case, and green the same case but 
		placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

@@

\newpage
\subsection{Non-trivial equilibrium configurations}
\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned 
with the electric field is relatively simple as the orientational 
torque is minimised (see Eq.\ref{eq:opt_torque}). Meaning once 
trapped the dimer is unlikely to change orientation enough to 
escape the trap. Regardless, that does not rule out the 
possibility that there is a stable configuration where the 
orientation not strictly vertical, in fact most experimental 
work with symmetric nano-dimers will trap them lying perpendicular 
to the beam direction \cite{Ahn2018, Reimann2018}. Unlike in Sec.~
\ref{sec:eq_positions} we cannot simply measure the optical force 
and torque as the parameter space is too large and determining if 
a particular position and orientation is stable is not clear based 
solely on force and torque measurements \cite{Bui2017}. Using the 
same simulation parameters as before we ran a number short 
simulations (total simulation time was $0.005$ s) with the laser 
power increased to $500\ mW$. Each simulation started with the 
dimer in a different starting position and orientation, due to the 
high laser power the dimers either escaped the trap or were stably 
trapped. The $z-\theta$ phase space - where $\theta$ is the angle 
between the direction of beam propagation and the dimer's orientation 
vector ($\theta=0^\circ$ is the 'standard' orientation) - can be 
divided into different regions depending on which equilibrium 
configuration is reached.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
		\caption{}
	\end{subfigure}
	%\captionsetup{hangindent=2.35cm}
	\caption{Map of $z-\theta$ phase space using a dimer of
		size ratio 2 with a laser power of 500 mW ($\theta=0
		\circ$ is the 'standard' orientation and $\theta=180
		^\circ$ is the 'inverted' orientation). The stable 
		configurations are indicated by the larger circles 
		and the starting conditions are colour coded to match 
		the stable point they end up in. Right hand render 
		shows a dimer in its off-axis configuration.}
	\label{fig:off_axis}
\end{figure}

Interestingly while the trap strength of these off-axis traps 
are similar in magnitude to the vertically aligned dimers, 
but when the laser power is lowered (around $5\ mW$) the traps 
become metastable; after reaching its equilibrium configuration 
the particle behaves similarly to a typically trapped dimer but 
the trapping potential is small enough that the dimer escapes 
in as little time as less than a second or after nearly a full 
3 seconds. Running similar simulations but for vertical 
configurations sees the dimer remaining trapped, even after 30 
seconds of run time, indicating that the trapping potential is 
far greater than the thermal energy. This suggests that the 
reason this off-axis configuration is metastable is due to the 
increased rotational freedom. Similar configurations have been 
explored with ellipsoids; Zhu \textit{et al.} looked at the 
dynamics of various elliptical particles and found that regardless 
of shape or initial orientation the particle would tend towards
either a purely vertical or purely horizontal orientation 
\cite{Zhu2021}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Continuous rotational motion in circularly polarised light}
One aspect that has yet to be covered in depth with regards 
to spherical aggregates of any construction is their interaction 
with circularly polarised light. For homogenous spheres the 
optical torque is regarded as being negligible as the spin 
density cannot impart angular momentum while propagating in a 
homogenous medium. Dimers however, have been shown to experience 
an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
while trapped in a circular polarised beam. In our simulations 
we found that dimers would rotate about their long axis when 
trapped in circularly polarised light. In this section we want 
to discuss how this behaviour is influenced by size, position, 
and orientation; and furthermore, we wish to address possible 
explanations for this behaviour, as none of the current
theories into optically induced rotation seem plausible.

\subsection{Polarisation Dependency on Dimer trajectory}
\label{sec:rot_pol}
In their paper Vigilante \textit{et al} attribute the rotational motion 
to the anisotropic shape of the dimer \cite{Vigilante2020}. Anisotropic 
scattering is a viable theory for describing optical rotation, however
there is usually an optical axis about which rotation occurs. For a dimer
this is typically the orientation vector that passes through the centres 
of both spheres \cite{Ahn2018, Reimann2018, Bruce2020}. There have been 
cases where the particle's cross sectional shape is engineered to scatter
light in one particular direction \cite{Higurashi1994}, but if this was
the case then the dimer should rotate when illuminated by any polarisation
of light.

To that end, we simulated the motion of an optically trapped dimer in 
beams of varying polarisation ($NA=1.2$, $P=100\ mW$), the dimer is 
composed of polystyrene ($n_p=1.59$, $n_m=1.33$). Each simulation was 
run for $1$ second ($\Delta t =10^-5$) and at the end we looked at the 
orientational time series; the dimer's orientation is recorded as a 
quaternion which can be easily converted to a 3-dimensional rotation 
matrix. By considering only the transverse components ($U_{1,x}$, 
$U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation matrix and taking 
the Fourier transformation of their time series reveals the rotational 
frequency. The laser power is set to $100\ mW$ to avoid large thermal 
fluctuations and so that the Fourier series of the transverse components 
approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function centred 
at the rotational frequency $\omega_{rot}$.

@@

This shows us that these optical rotations are polarisation 
dependent and not merely an example of the dimer scattering 
light asymmetrically. We also know that this cannot be a 
similar mechanism to that of experimental works \cite{Ahn2018, 
Reimann2018} as in our simulations the dimer's long axis
is parallel to the electric field. In this orientation 
the dimer's optical axis is aligned perpendicular to the
polarisation vector meaning that there should be no angular
momentum transferred along the optical axis. This is consistent 
with other experiments involving elongated particles. A dual
beam trap was used to study the dynamics of "disk like" particles. 
They found that these particles had one of two stable
orientations: Either orienting with the 'flat' side perpendicular
to the beam trap, with no rotation being observed \cite{Brzobohaty2015}. 
Or orientating with the 'flat' side parallel to the trap, in which 
case the long axis was aligned with the polarisation vector and 
thus rotational motion was detected \cite{Brzobohaty2015}.

\subsection{Brownian Vortex via Curl of Spin momentum}
In their work Vigilante refers to the spin-curl effects 
demonstrated by Grier \textit{et al} \cite{Ruffner2012}, 
in which the curl of the spin density leads to a second 
order optical force that orbits around the beams central 
axis \cite{Yevick2017}. While several papers have 
demonstrated this phenomena \cite{Zhao2007, Zhao2009, 
Wang2010} it was only properly formalised by \cite{Ruffner2012}. 
In which they showed that the seemingly random trajectory 
of a trapped sphere was biased by the polarisation state 
of the trapping beam. While not immediately evident from 
the trajectory the helicity of the trapping beam was revealed 
by computing the particle's probability flux using.
\begin{align}
	j(r) = \frac{1}{N-1} \sum_{j=1}^{N-1}
	\frac{r_{j+1}-r_j}{\tau}\delta_{sigma_j}\left(r-\frac{r_{j+1}+r_j}{2}\right)
	\label{eq:prob_flux}
\end{align}

@@

While the results from \cite{Ruffner2012} suggest that the 
optical rotation seen in asymmetric dimers can be attributed to 
the same spin-curl forces there are several questions that 
cannot be explained purely by the spin-curl force.

\subsection{Optical torque differences}
In their paper Grier \textit{et al} show that a 7 sphere cluster
will rotate when trapped in circularly polarised light but a 
homogenous sphere will have a slight curl to its trajectory. 
They attribute this to the curl of the spin angular momentum
generating a "Brownian vortex". But what is more interesting, 
and something that appears unaddressed, is how the effects of 
this vortex change based on the overall shape of the target
particle.

You would not expect that the addition of a single additional 
sphere should drastically adjust the torque especially if said 
sphere is relatively small. However when we measured the optical 
torque of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we 
found the exact opposite. In both cases we used the same trapping 
beam as used for fig.~\ref{fig:rotation_vs_pol} but with a 
circularly polarised beam. Both the sphere and dimer were rotated 
in the $x-z$ plane and the all three components of the optical 
torque were recorded.
\begin{figure}[h!]
	\centering	
	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
	\caption{Optical toque experienced by a dimer ($a_{I}/a
	{II}=10$) and a single isotropic sphere. Both were rotated 
	in the $x-z$ plane and the angle between $U_z$ and the beam 
	axis gives the polar angle.The solid lines denote the torque 
	experienced by the dimer whereas the dashed lines represent 
	the torque experienced by the sphere.}
	\label{fig:sphere_dimer_torque}
\end{figure}

The torques about the $x$ and $y$ axis can be somewhat understood 
as the second sphere is being drawn back towards the centre of the 
trap by the gradient forces. The same cannot be said for the $z$-
component of the optical torque, which is non-zero even at a 
$80^\circ$ angle. This cannot be simply explained via spin-curl 
effects, as the 'Brownian vortex' should instead be driving the
dimer around the beam axis - similar to how the 7 sphere cluster 
behaved. What is clear that the mere combination of two spheres
results in a unique behaviour that has not been previously 
investigated.

\subsection{Rotational frequency as a function of size ratio and 
	orientation}
\label{sec:rot_size}	
One factor t Intuitively, you would expect that a 
larger particle would experience a greater torque and therefore 
rotate faster. By repeating the same kinds of simulation as used 
in \ref{fig:rotation_vs_pol} but for a circularly polarised beam 
$\phi=90\circ$ it was found that not only is the rotation rate 
dependent on the size of the dimer, but also on its orientation 
and therefore their axial position.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
  \caption{Rotation rate plotted against dimer size ratio while 
  	trapped in a circularly polarised beam; a positive rotation 
  	rate indicates clockwise rotation, whereas a negative 
  	rotation rate indicates counter-clockwise rotation. The red 
  	line is for the case of a dimer in its 'standard' orientation. 
  	The blue line is for the case when the dimer is in its 
  	'inverted' orientation while trapped above the focus of the 
  	beam. And lastly the green line is for the case when the 
  	dimer is in its 'inverted' orientation, but when it is trapped 
  	below the focus of the beam.}
\end{figure}

It is difficult to see from the graph, but the rotation rate never
truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which 
would imply that a second sphere of radius $200\,{\rm nm}$ is enough 
to induce rotational motion. What is also interesting is that the 
rotation rate is not correlated directly with either the particle 
size or the equilibrium position. This is in stark contrast to previous
reports of dimer optical rotation; Ahn \textit{et al.} reported that
the rotational frequency is maximised when the dimer is symmetric 
\cite{Ahn2018}. Their work was conducted in a vacuum so the only 
limiting factor to the optical rotational motion is the structural 
stability of the particle itself.

\subsection{Gyroscopic Precession using asymmetric dimers}
As mentioned in section~\ref{sec:off-axis} for specificity sized 
dimers there is the potential for non-vertical trapping orientations 
in which the dimer is still stably trapped. When a circularly 
polarised beam is used the dimer exhibits gyroscopic precession. 
As shown in fig.~\ref{fig:gyro} the dimer's trajectory exhibits 
periodic rotation, both around its long axis and about the beam 
axis.
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped 
		in an off axis orientation with a circularly polarised beam 
		($P= 100\ mW$). The far left column depicts the dimer's centre 
		of mass position with time; the remaining 3 columns show the 9
		components of the dimers' rotation matrix, with each column being
		associated with one of its three principal axis. }
	\label{fig:gyro}
\end{figure}

Applying a Fourier analysis to the above trajectory reveals 
the 3 fundamental frequencies typically associated with 
gyroscopic precession. Fig.~\ref{fig:gyro_diagram} gives 
a representative idea of the motion seen in fig.~\ref{fig:gyro}.
Firstly there is precession, is denoted by $\psi$, this
causes the dimer to rotate in the $XY$ plane while maintaining
a constant polar angle. The precession can be due to the dimer 
having inhomogeneous polarisation susceptibility. Since its 
long axis is more susceptible to being polarised the dimer will
try to align its long axis with the polarisation vector 
\cite{Bruce2020}. Second there is nutation, denoted by $\theta$, 
which is due to the dimer's Brownian motion being influenced by 
the 'Brownian vortex', as predicted by \cite{Ruffner2012}. And 
lastly there is rotation, denoted by $\omega$, where the dimer
spins around its long axis as it did in \ref{sec:rot_pol} \&
\ref{sec:rot_size}.

What is interesting is that while we see the precession and 
nutation also contribute to the rotation frequency. As can be 
seen in both fig.~\ref{fig:gyro} and fig.~\ref{fig:gyro_diagram}
the rotation matrix components $U_{x,x}$, $U_{x,y}$, $U_{y,x}$
and $U_{y,y}$

\begin{figure}
	\hspace{-2.3cm}
	\includegraphics[width=1.2\linewidth]{gyroscopic_diagram.pdf}
	\caption{Representative diagram of the gyroscopic precession 
	from fig.~\ref{fig:gyro}. The dimer has three principal angular
	frequencies: The rotation ($\omega$) occurs around the 
	the dimer's long axis. Precession ($\psi$) is seen where the dimer 
	rotates around the beam axis. Nutation ($\theta$) is due to the dimer's 
	centre of diffusion orbiting the beam axis. Shown on the right
	is the power spectra from fig.~\ref{fig:gyro} with the associated 
	frequencies labelled. The power spectrum have been zoomed in on
	the relevant frequencies to highlight the precise values.}
	\label{fig:gyro_diagram}
\end{figure}

This gyroscopic motion has been demonstrated previously in 
nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
but has not been observed for micron scale aggregates. What
is interesting that the precession is seen around the particle's
optical axis (in the case of a dimer this would be its long 
axis). Fig.~\ref{fig:gyro} instead shows the precession 
occurring perpendicular to the optical axis. In the former 
case the precession can be explained as simply occurring due 
to the reaction torque from the surrounding fluid \cite{Zhu2021}.
In their analysis of ellipsoidal particles, Zhu \textit{et al.}
found that when the particle rotated around their minimal inertial 
axis (similar to the observed behaviour in dimers), the particle 
would quickly 'stabilise' by aligning its' optical axis with the 
beam axis. This highlights that a spherical dimer is capable of 
rotating about its minimal inertial axis while remaining stably
trapped. This has potential implications beyond simply 
understanding particle dynamics.
